---
title: 'Problem Set #4'
author: "Ruoqi Zhang"
date: "2/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(readxl)
library(lubridate)
library(ggridges)
library(janitor)
library(stringr)
library(gt)
library(scales)
```

```{r data_preparation, echo=FALSE, message=FALSE}
poll <- read_csv("ps_4_elections-poll-nc09-3.csv")

# poll_clean <- poll_original %>% 
#   anti_join(filter_all(poll_original, any_vars(str_detect(., pattern = "DO NOT READ"))))
# 
#
# poll_invalid <- poll_original %>% 
#   filter_all(any_vars(str_detect(., pattern = "DO NOT READ")))
# poll_clean <- anti_join(poll_original, poll_invalid)
# 
# poll_clean <- poll_original[!(poll_original$timestamp %in% poll_invalid$timestamp),]
# 
# poll_clean <- poll_original %>% 
#   filter_all(any_vars(str_detect(., pattern = "DO NOT READ"))) %>% 
#   anti_join(poll_original)
#   
#   
#   filter_all(all_vars(!grepl('[DO NOT READ]',.)))
# 
# poll_clean <- poll_original %>% 
#   filter(poll)

```

# Question 1

```{r q1_1, echo=FALSE}
ans_q1_1 <- poll %>% 
  filter(response=="Dem") %>% 
  count()
```

There were `r ans_q1_1` respondents who supported the Democratic candidate.

```{r q1_2, echo=FALSE}
rep_num <- poll %>% 
  filter(response == "Rep") %>% 
  count()

und_num <- poll %>% 
  filter(response %in% c('Und')) %>% 
  count()

ans_q1_2 <- rep_num - und_num 
```

There were `r ans_q1_2` more respondents who favored the Republican candidate than who were Undecided.

```{r q1_3, echo=FALSE}
ans_q1_3 <- poll %>% 
  filter(gender != gender_combined) %>% 
  count()
```

There are two gender variables (`gender` and `gender_combined`). There are `r ans_q1_3` individuals for whom these variables have different values.

```{r q1_4, echo=FALSE}
ans_q1_4 <- poll %>% 
  filter(race_eth == "White",
         race_eth != file_race_black) %>% 
  count()
```

There are `r ans_q1_4` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

```{r q1_5, echo=FALSE}
rep_earliest <- poll %>% 
  filter(response == "Rep") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)

dem_earliest <- poll %>% 
  filter(response == "Dem") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)

ans_q1_5 <- round(as.numeric(dem_earliest - rep_earliest))

```

The first response of Rep came `r ans_q1_5` minutes (rounded to the nearest minute) before the first `response` of Dem.

# Question 2

```{r table, echo=FALSE}

## TODO: Replace the undecided cell into "-"

poll %>% 
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  select(response, race_eth, final_weight) %>% 
  group_by(response, race_eth) %>% 
  summarize(total = sum(final_weight)) %>% 
  spread(key = response, value = total) %>% 
  mutate_all(list(~replace_na(.,0))) %>% 
  mutate(total_weight = Dem + Rep + Und + `3`) %>% 
  mutate_all(list(~na_if(.,0))) %>% 
  mutate(Dem = percent(Dem/total_weight, 1),
         Rep = percent(Rep/total_weight, 1),
         Und = percent(Und/total_weight, 1)) %>% 
  select(race_eth, Dem, Rep, Und) %>% 
  mutate(race_eth = factor(race_eth, levels = c("White", "Black", "Hispanic", "Asian", "Other"))) %>% 
  arrange(race_eth) %>% 
  mutate_all(list(~str_replace(.,"NA%","-"))) %>% 
  gt() %>% 
  tab_header(
    title = "Polling results in North Carolina 9th Congressional District"
  ) %>% 
  tab_stubhead_label(label = "race_eth") %>% 
  cols_label(
    race_eth = "Race/Ethnicity",
    Dem = "DEM.",
    Rep = "REP.",
    Und = "UND."
  ) %>% 
  tab_source_note(
    source_note = "Source: New York Times Upshot/Siena College 2018 live polls"
  )

```

# Question 3

```{r q3, echo=FALSE}
poll %>% 
  filter(educ != "[DO NOT READ] Refused") %>% 
  select(educ, final_weight) %>% 
  mutate(educ = factor(educ, levels = c("Grade school", "High school",
                                        "Some college or trade school", "Bachelors' degree",
                                        "Graduate or Professional Degree"))) %>% 
  arrange(educ) %>% 
  ggplot(aes(x=educ, y=final_weight)) + 
    geom_violin() + 
    geom_jitter(alpha = 0.25, width = 0.25) + 
    coord_flip() +
    labs(title = "More Educated Matter Less in North Carolina 9th",
           subtitle = "Poll gives more weight to people who are less likely to participate in polls",
         x = NULL,
         y = "Weight Given to Respondent in Calculating Poll Results",
           caption = "New York Times Upshot/Siena College 2018 live polls")
```

# Question 4

# Collaborators
Neil Khruana